---
subtitle: "Exploring Distributional Shapes"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{polyPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 7, comment = NA, fig.path = "figures/")
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r,include=FALSE}
if (!require(polyPlot)) {
  if (!require(remotes)) install.packages("remotes")
  remotes::install_github("cwendorf/polyPlot")
}
library(polyPlot)
```

## Exploring Distributional Shapes

The examples below demonstrate how the PolyPlot does a good job of approximating the shape of the underlying distribution, even in cases of skewed data.

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile, base_level = 3, toc_depth = 4)
```

---

### Positively Skewed Data

For example, simulate some positively skewed data. Use `polyPlot` to get the statistics plot and a frequency histogram. The calculated values are not presented on the plot, but are given in a separate table.
```{r, Shapes-Positive}
Scores <- round(rbeta(100, 2, 5) * 20, 0)
polyStats(Scores, digits = 3)
polyPlot(Scores, values = FALSE, type = "frequency", main = "PolyPlot and Histogram for Positively Skewed Data")
```

### Negatively Skewed Data

As another example, simulate some negatively skewed data. Use `polyPlot` to get the statistics plot and a density curve. The calculated values are not presented on the plot, but are given in a separate table.
```{r, Shapes-Negative}
Scores <- round(rbeta(100, 5, 2) * 20, 0)
polyStats(Scores, digits = 3)
polyPlot(Scores, values = FALSE, type = "density", main = "PolyPlot and Density for Negatively Skewed Data")
```

### Bimodal Data

As a final example, simulate some bimodal data. Here the values are presented on each `polyPlot` to show their utility in explaining the underlying data.
```{r, Shapes-Bimodal}
Scores <- c(round(rnorm(50, 7, 1), 0), round(rnorm(50, 13, 1), 0))
polyPlot(Scores, type = "frequency", col = "blue", bg = "goldenrod", values = TRUE, main = "PolyPlot and Histogram for Bimodal Data")
polyPlot(Scores, type = "density", col = "blue", bg = "goldenrod", values = TRUE, main = "PolyPlot and Density for Bimodal Data")
```
