---
subtitle: "Basic Tutorial for polyPlot"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Basic Tutorial for polyPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
# suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width = 7, fig.height = 5, comments = NA)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r,include=FALSE}
if (!require(polyPlot)) {
  if (!require(remotes)) install.packages("remotes")
  remotes::install_github("cwendorf/polyPlot")
}
library(polyPlot)
```

## Basic Tutorial for polyPlot

The following codes demonstrates the default options of PolyPlot. It also explains the basic statistics used.

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile, base_level = 3, toc_depth = 4)
```

---

### The PolyPlot

First, simulate a set of 100 scores from a normal distribution. Then call `polyPlot` to get the plot.
```{r,Basic-Default}
Scores <- round(rnorm(100, 10, 2), 0)
polyPlot(Scores)
```

Generally speaking, each level/row provides a different category of measures of location and spread:

- First/Top: Provides the quartiles (and median as the center) of the distribution, with the interquartile range divided by 2 as the measure of spread
- Second: Provides the means of the halves (and the whole set as the center) of the distribution, with the mean absolute deviation from the median as the measure of spread
- Third: Provides the mean (as the center) of the distribution and points +/- one standard deviation from the mean (i.e, z scores of -1 and 1), with the standard deviation as the measure of spread
- Fourth/Bottom: Provides the range (and midrange as the center) of the distribution, with the range divided by 2 as the measure of spread

### Adding Background Plots

To see the relationship between the PolyPlot and the underlying distribution, either a frequency distribution (histogram) or a density curve can be added in the background.
```{r,Basic-Background}
polyPlot(Scores, type = "frequency")
polyPlot(Scores, type = "density")
```

Further explanation of the math behind the PolyPlot is available in Seier and Bonett (2011).
