---
title: "PolyPlot2"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: tango
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Basic Tutorial for PolyPlot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
if(!require(polyPlot2)) {
if(!require(remotes)) install.packages('remotes')
remotes::install_github("cwendorf/polyPlot2")}
library(polyPlot2)
```

## Basic Tutorial for PolyPlot2

The following codes demonstrates the default options of PolyPlot. It also explains the basic statistics used.

### The PolyPlot

First, simulate a set of 100 scores from a normal distribution. Then call `polyPlot` to get the plot.
```{r,Basic-Default}
Scores <- round(rnorm(100,10,2),0)
polyPlot(Scores)
```

Generally speaking, each level/row provides a different category of measures of location and spread:

- First/Top: Provides the quartiles (and median as the center) of the distribution, with the interquartile range divided by 2 as the measure of spread
- Second: Provides the means of the halves (and the whole set as the center) of the distribution, with the mean absolute deviation from the median as the measure of spread
- Third: Provides the mean (as the center) of the distribution and points +/- one standard deviation from the mean (i.e, z scores of -1 and 1), with the standard deviation as the measure of spread
- Fourth/Bottom: Provides the range (and midrange as the center) of the distribution, with the range divided by 2 as the measure of spread

### Adding Background Plots

To see the relationship between the PolyPlot and the underlying distribution, either a frequency distribution (histogram) or a density curve can be added in the background.
```{r,Basic-Background}
polyPlot(Scores,type="frequency")
polyPlot(Scores,type="density")
```

Further explanation of the math behind the PolyPlot is available in Seier and Bonett (2011).
