---
title: "PolyPlot2"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: tango
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Advanced Tutorial for PolyPlot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
if(!require(polyPlot2)) {
if(!require(remotes)) install.packages('remotes')
remotes::install_github("cwendorf/polyPlot2")}
library(polyPlot2)
```

## Advanced Tutorial for PolyPlot2

The following code demonstrates some of the available options for customizing the PolyPlot. It also demonstrates how the PolyPlot does a good job of approximating the shape of the underlying distribution.

### The PolyPlot for Positively Skewed Data

Simulate some positively skewed data. Then use `polyStats` get the values separate from the `polyPlot` (which in this case features a frequency histogram).
```{r<advanced-Positive}
Scores <- round(rbeta(100,2,5)*20,0)
polyStats(Scores,digits=3)
polyPlot(Scores,type="frequency",col="blue",bg="goldenrod",values=FALSE,main="PolyPlot and Histogram for Positively Skewed Data")
```

### The PolyPlot for Negatively Skewed Data

Simulate some negatively skewed data. Then use `polyStats` get the values separate from the `polyPlot` (which in this case features a density curve).
```{r,Advanced-Negative}
Scores <- round(rbeta(100,5,2)*20,0)
polyStats(Scores,digits=3)
polyPlot(Scores,type="density",col="blue",bg="goldenrod",values=FALSE,main="PolyPlot and Density for Negatively Skewed Data")
```

### The PolyPlot for Bimodal Data

Simulate some bimodal data. The values are presented on each `polyPlot` to show their utility in explaining the underlying data.
```{r,Advanced-Bimodal}
Scores <- c(round(rnorm(50,7,1),0),round(rnorm(50,13,1),0))
polyPlot(Scores,type="frequency",col="blue",bg="goldenrod",values=TRUE,main="PolyPlot and Histogram for Bimodal Data")
polyPlot(Scores,type="density",col="blue",bg="goldenrod",values=TRUE,main="PolyPlot and Density for Bimodal Data")
```
