---
title: "PolyPlot2"
author: "Craig A. Wendorf"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    collapse: true
    theme: cerulean
    highlight: tango
    keep_md: TRUE
vignette: >
  %\VignetteIndexEntry{Tutorial for PolyPlot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
#suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width=7, fig.height=5,fig.path='figures/',echo=TRUE,warning=FALSE,message=FALSE)
```

```{r,include=FALSE}
library(polyplot2)
```

## Tutorial for PolyPlot2

### The Basic PolyPlot

In order to demonstrate the finer points of a PolyPlot, the following code simulates 1000 scores from a positively skewed unimodal distribution and then calls `polyplot2` using the default options.

```{r}
Scores <- c(rnorm(n=600,mean=100,sd=15),rnorm(n=200,mean=115,sd=20),rnorm(n=200,mean=130,sd=25))
polyplot2(Scores)
```

By default, the function labels all of the points and provides summary statistics in the margin. Generally speaking, each level/row provides a different category of measures of location and spread.

- First/Top: Provides the mode of the distribution, with the sample size listed in the right margin
- Second: Provides the quartiles (and median) of the distribution, with the interquartile range divided by 2 in the right margin
- Third: Provides the means of the halves (and the whole) of the distribution, with the mean absolute deviation from the median in the right margin
- Fourth: Provides the mean of the distribution and points +/- one standard deviation from the mean, with the standard deviation in the right margin
- Fifth/Bottom: Provides the range (and midrange) of the distribution, with the range divided by 2 in the right margin

### Using the PolyPlot to Determine Shape

The following code demonstrates some of the available options for customizing the PolyPlot.Similarly, it demonstrates how the PolyPlot does a good job of approximating the shape of the underlying distribution.

```{r}
par(mfrow=c(1,2))
polyplot2(Scores,plot="frequency",type="bar",values=FALSE,main="PolyPlot2 plus Frequency Histogram",col="darkblue",bg="darkgoldenrod")
polyplot2(Scores,plot="density",type="curve",values=FALSE,main="PolyPlot2 plus Density Curve",col="darkblue",bg="darkgoldenrod")
```

Further explanation of the polyplot and how it can be used to understand the shape of the distribution is available in Seier and Bonett (2011).
