---
subtitle: "Advanced Tutorial for polyPlot"
output:
  github_document:
    preserve_yaml: FALSE
vignette: >
  %\VignetteIndexEntry{Advanced Tutorial for polyPlot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
# suppress the warnings and other messages from showing in the knitted file.
knitr::opts_chunk$set(fig.width = 7, fig.height = 6, fig.path = "figures/", echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
devtools::source_gist("8e6e5dc401e3fc1042ef7a030f9d19c7", filename = "revised_toc.R")
```

```{r,include=FALSE}
if (!require(polyPlot)) {
  if (!require(remotes)) install.packages("remotes")
  remotes::install_github("cwendorf/polyPlot")
}
library(polyPlot)
```

## Advanced Tutorial for polyPlot

The following code demonstrates some of the available options for customizing the PolyPlot. It also demonstrates how the PolyPlot does a good job of approximating the shape of the underlying distribution.

```{r toc, echo=FALSE}
thisfile <- knitr::current_input()
revised_toc(thisfile, base_level = 3, toc_depth = 4)
```

---

### The PolyPlot for Positively Skewed Data

Simulate some positively skewed data. Then use `polyStats` get the values separate from the `polyPlot` (which in this case features a frequency histogram).
```{r,Advanced-Positive}
Scores <- round(rbeta(100, 2, 5) * 20, 0)
polyStats(Scores, digits = 3)
polyPlot(Scores, type = "frequency", col = "blue", bg = "goldenrod", values = FALSE, main = "PolyPlot and Histogram for Positively Skewed Data")
```

### The PolyPlot for Negatively Skewed Data

Simulate some negatively skewed data. Then use `polyStats` get the values separate from the `polyPlot` (which in this case features a density curve).
```{r,Advanced-Negative}
Scores <- round(rbeta(100, 5, 2) * 20, 0)
polyStats(Scores, digits = 3)
polyPlot(Scores, type = "density", col = "blue", bg = "goldenrod", values = FALSE, main = "PolyPlot and Density for Negatively Skewed Data")
```

### The PolyPlot for Bimodal Data

Simulate some bimodal data. The values are presented on each `polyPlot` to show their utility in explaining the underlying data.
```{r,Advanced-Bimodal}
Scores <- c(round(rnorm(50, 7, 1), 0), round(rnorm(50, 13, 1), 0))
polyPlot(Scores, type = "frequency", col = "blue", bg = "goldenrod", values = TRUE, main = "PolyPlot and Histogram for Bimodal Data")
polyPlot(Scores, type = "density", col = "blue", bg = "goldenrod", values = TRUE, main = "PolyPlot and Density for Bimodal Data")
```
